---
title: "G-Ranges"
output: html_notebook
---

#1- open libraries
```{r}
library(tidyverse)
library(readr)
library(GenomicRanges)
```


#2- Open CSV file
```{r}
BC01_Nucleosomes <- read_delim("Nucleosome_Widom_redefined.csv")
head(BC01_Nucleosomes)
```
```{r}
GR_BC01 <- BC01_Nucleosomes %>% 
  select(ReadName, Mapping_y, MapStart_y, MapEnd_y, StrandStatus, StrandDirection) %>% 
  mutate(StrandRep = if_else(StrandStatus=="Lead","+","-"))
GR_BC01
```

```{r}
GR_BC01_format <- GRanges(seqnames = GR_BC01$Mapping_y,
        ranges = IRanges(start = GR_BC01$MapStart_y, end = GR_BC01$MapEnd_y),
        strand = GR_BC01$StrandRep,
        strand_status = GR_BC01$StrandStatus)

GR_BC01_format
  
```
#Make the coverage. In this case is the total.
```{r}
Cover <- coverage(GR_BC01_format)
as.numeric(Cover$pFB2)
```
#Preparing data in better format
```{r}
df_BC01_orig <-data.frame("Position" = seq_along(as.numeric(Cover$pFB2)), 
                          "Coverage" = as.numeric(Cover$pFB2), Type_coverage = "barcode01")
head(df_BC01_orig)
```

Plot the coverage
```{r}
ggplot(df_BC01_orig) +
  geom_area(aes(x = Position, y = Coverage))
  
```
#Separate for Leading and then lagging.
```{r}
GR_BC01_Leading <- GR_BC01 %>% 
  filter(StrandStatus == "Lead")
GR_BC01_Leading

GR_BC01_Lagging <- GR_BC01 %>% 
  filter(StrandStatus == "Lag")
GR_BC01_Lagging


```
#Transform to Granges
```{r}
GR_BC01_format_Leading <- GRanges(seqnames = GR_BC01_Leading$Mapping_y,
        ranges = IRanges(start = GR_BC01_Leading$MapStart_y, end = GR_BC01_Leading$MapEnd_y),
        strand = GR_BC01_Leading$StrandRep,
        strand_status = GR_BC01_Leading$StrandStatus)
GR_BC01_format_Leading

GR_BC01_format_Lagging <- GRanges(seqnames = GR_BC01_Lagging$Mapping_y,
        ranges = IRanges(start = GR_BC01_Lagging$MapStart_y, end = GR_BC01_Lagging$MapEnd_y),
        strand = GR_BC01_Lagging$StrandRep,
        strand_status = GR_BC01_Lagging$StrandStatus)
GR_BC01_format_Lagging

```
#Make coverage for Leading
```{r}
Cover_leading <- coverage(GR_BC01_format_Leading)
as.numeric(Cover_leading$pFB2)
```

#Make covergae for Lagging
```{r}
Cover_lagging <- coverage(GR_BC01_format_Lagging)
as.numeric(Cover_lagging$pFB2)
```

#Prepare data for a better format
```{r}
df_BC01_leading <-data.frame("Position" = seq_along(as.numeric(Cover_leading$pFB2)), 
                          "Coverage" = as.numeric(Cover_leading$pFB2), Type_coverage = "barcode01")
head(df_BC01_leading)

df_BC01_lagging <-data.frame("Position" = seq_along(as.numeric(Cover_lagging$pFB2)), 
                          "Coverage" = as.numeric(Cover_lagging$pFB2), Type_coverage = "barcode01")
head(df_BC01_lagging)
```
Plot the coverages. Leading
```{r}
ggplot(df_BC01_leading) +
  geom_area(aes(x = Position, y = Coverage))
```
Plot the coverages. Lagging

```{r}
ggplot(df_BC01_lagging) +
  geom_area(aes(x = Position, y = Coverage))
```
#Separate in the 4 different quadrants. Maybe it is easy to plot.
```{r}
head(GR_BC01)
unique(GR_BC01$StrandDirection)

GR_BC01_Left_Lag <- GR_BC01 %>% 
  filter(StrandDirection == "fwdLag")
GR_BC01_Left_Lag

GR_BC01_Left_Lead <- GR_BC01 %>% 
  filter(StrandDirection == "revLead")
GR_BC01_Left_Lead

GR_BC01_Right_Lag <- GR_BC01 %>% 
  filter(StrandDirection == "revLag")
GR_BC01_Right_Lag

GR_BC01_Right_Lead <- GR_BC01 %>% 
  filter(StrandDirection == "fwdLead")
GR_BC01_Right_Lead

```
 
 #Transform to GRanges
```{r}
GR_BC01_format_left_lagging <- GRanges(seqnames = GR_BC01_Left_Lag$Mapping_y,
        ranges = IRanges(start = GR_BC01_Left_Lag$MapStart_y, end = GR_BC01_Left_Lag$MapEnd_y),
        strand = GR_BC01_Left_Lag$StrandRep,
        strand_status = GR_BC01_Left_Lag$StrandStatus,
        strand_direction = GR_BC01_Left_Lag$StrandDirection)
GR_BC01_format_left_lagging

GR_BC01_format_left_leading <- GRanges(seqnames = GR_BC01_Left_Lead$Mapping_y,
        ranges = IRanges(start = GR_BC01_Left_Lead$MapStart_y, end = GR_BC01_Left_Lead$MapEnd_y),
        strand = GR_BC01_Left_Lead$StrandRep,
        strand_status = GR_BC01_Left_Lead$StrandStatus,
        strand_direction = GR_BC01_Left_Lead$StrandDirection)
GR_BC01_format_left_leading

GR_BC01_format_Right_lagging <- GRanges(seqnames = GR_BC01_Right_Lag$Mapping_y,
        ranges = IRanges(start = GR_BC01_Right_Lag$MapStart_y, end = GR_BC01_Right_Lag$MapEnd_y),
        strand = GR_BC01_Right_Lag$StrandRep,
        strand_status = GR_BC01_Right_Lag$StrandStatus,
        strand_direction = GR_BC01_Right_Lag$StrandDirection)
GR_BC01_format_Right_lagging

GR_BC01_format_Right_leading <- GRanges(seqnames = GR_BC01_Right_Lead$Mapping_y,
        ranges = IRanges(start = GR_BC01_Right_Lead$MapStart_y, end = GR_BC01_Right_Lead$MapEnd_y),
        strand = GR_BC01_Right_Lead$StrandRep,
        strand_status = GR_BC01_Right_Lead$StrandStatus,
        strand_direction = GR_BC01_Right_Lead$StrandDirection)
GR_BC01_format_Right_leading

```
```{r}
Cover_left_lag <- coverage(GR_BC01_format_left_lagging)
as.numeric(Cover_left_lag$pFB2)

Cover_left_lead <- coverage(GR_BC01_format_left_leading)
as.numeric(Cover_left_lead$pFB2)

Cover_rigth_lag <- coverage(GR_BC01_format_Right_lagging)
as.numeric(Cover_rigth_lag$pFB2)

Cover_rigth_lead <- coverage(GR_BC01_format_Right_leading)
as.numeric(Cover_rigth_lead$pFB2)

```
 #Prepare data for better format
```{r}
df_BC01_Left_lag <-data.frame("Position" = seq_along(as.numeric(Cover_left_lag$pFB2)), 
                          "Coverage" = as.numeric(Cover_left_lag$pFB2), Type_coverage = "Left_lag")
head(df_BC01_Left_lag)

df_BC01_Left_lead <-data.frame("Position" = seq_along(as.numeric(Cover_left_lead$pFB2)), 
                          "Coverage" = as.numeric(Cover_left_lead$pFB2), Type_coverage = "Left_lead")
head(df_BC01_Left_lead)

df_BC01_Rigth_lag <-data.frame("Position" = seq_along(as.numeric(Cover_rigth_lag$pFB2)), 
                          "Coverage" = as.numeric(Cover_rigth_lag$pFB2), Type_coverage = "Rigth_lag")
head(df_BC01_Rigth_lag)

df_BC01_Right_lead <-data.frame("Position" = seq_along(as.numeric(Cover_rigth_lead$pFB2)), 
                          "Coverage" = as.numeric(Cover_rigth_lead$pFB2), Type_coverage = "Right_lead")
head(df_BC01_Right_lead)
```
#Plot
```{r}
ggplot(df_BC01_Left_lag) +
  geom_area(aes(x = Position, y = Coverage))

ggplot(df_BC01_Left_lead) +
  geom_area(aes(x = Position, y = Coverage))


ggplot(df_BC01_Rigth_lag) +
  geom_area(aes(x = Position, y = Coverage))

ggplot(df_BC01_Right_lead) +
  geom_area(aes(x = Position, y = Coverage))
```
#Join the dataframes in order to plot all of them.

```{r}
plot_combine_df <- rbind((df_BC01_Left_lag),(df_BC01_Left_lead),(df_BC01_Rigth_lag),(df_BC01_Right_lead))
unique(plot_combine_df$Type_coverage)
```
#Make the plot of all qudrants
```{r}
ggplot(plot_combine_df) +
  geom_area(aes(x = Position, y = Coverage)) +
  facet_grid(rows = vars(Type_coverage))
```


 
