---
title: "N-Sonar data analysis"
output: html_notebook
---

#1- open libraries
```{r}
library(tidyverse)
library(readr)
```

#2- Open CSV file
```{r}
BC01_Nucleosomes <- read_delim("Nucleosome_Widom_redefined.csv")
head(BC01_Nucleosomes)
```
#3- Start to disecting the tibble by filter the required data:
A) Confirm total number of nucleosomes

```{r}
T_n <- BC01_Nucleosomes %>%
  nrow()
T_n
```

B) Number of nucleosomes in leading strands. Should be only reads mapped to the plasmid reference.

Fist check the values in the Mapaping_y and StrandStatus columns
```{r}
unique(BC01_Nucleosomes$Mapping_y)
unique(BC01_Nucleosomes$StrandStatus)
```


```{r}
T_n_Leading <- BC01_Nucleosomes %>% 
  filter(StrandStatus == "Lead") %>%
  nrow()
T_n_Leading
```

C) Number of nucleosomes in lagging strands. Should be only reads mapped to the plasmid reference.

```{r}
T_n_lagging <- BC01_Nucleosomes %>%
  filter(StrandStatus == "Lag") %>% 
  nrow()
T_n_lagging 

```
D) Calculate the % of nucleosomes 
```{r}
T_n_Leadlag <-BC01_Nucleosomes %>% 
  count(StrandStatus) %>% 
  mutate(percentage = n/T_n * 100) %>% 
  mutate(StrandStatus = factor(StrandStatus, levels = c("Lead", "Lag"))) %>% 
  complete(StrandStatus, fill = list(n = 0, percentage = 0 ))
T_n_Leadlag
```
#Ploting the barchart
```{r}
p1 <- ggplot(T_n_Leadlag, aes(x = StrandStatus, y= percentage, fill = StrandStatus)) +
  geom_bar(
    stat = "identity", 
    color = "black",
    size = 0.25,
    width =0.8) +
  scale_y_continuous(
    breaks = seq(0, 100, by =10), 
    limits = c(0, 100), 
    expand = c(0,0)) +
  geom_text(
    aes(label = paste0(round(percentage, 1), "%")),
            vjust = -0.5) +
  geom_text(
    aes(label = n),
    vjust = 1.5) +
  labs(
    y= "% of Nucleosomes", 
    x = "Strands") +
  scale_x_discrete(
    labels = c("Leading", "Lagging")) +
  theme(
    legend.position = "none",
    axis.line = element_line(color = "black", size = 0.5),
    axis.text.x = element_text(color = "black"),
    axis.text.y = element_text(color = "black"),
    plot.title = element_text(hjust = 0.5),
    panel.background = element_blank(),
    plot.background = element_blank()) +
  ggtitle("Total nucleosomes")

  ggsave("Total_nucleosomes_BC01.svg", plot = p1)

print(p1)
```
E) Calculating nucleosomes in right and left fork
```{r}
unique(BC01_Nucleosomes$StrandDirection)
```
#Calculating percentage for each strand in the right and left fork
```{r}
Fork_nucleosomes <- BC01_Nucleosomes %>% 
  count(StrandStatus, StrandDirection) %>% 
  mutate(percentage = n/T_n * 100) %>% 
  mutate(StrandDirection = factor(StrandDirection, levels = c("revLead", "fwdLag", "fwdLead", "revLag"))) %>% 
  complete(StrandStatus, fill = list(n = 0, percentage = 0))
Fork_nucleosomes
```
#Ploting the barchart
```{r}
p2 <- ggplot(Fork_nucleosomes, aes(x = StrandDirection, y = percentage, fill = rev(StrandStatus))) +
  geom_col(
    stat = "identity",
     color = "black",
    size = 0.25,
    width = 0.8) +
  scale_y_continuous(
    breaks = seq(0, 500, by =10), 
    limits = c(0, 50), 
    expand = c(0,0)) +
  geom_text(
    aes(label = paste0(round(percentage, 1), "%")),
            vjust = -0.5) +
  geom_text(
    aes(label = n),
    vjust = 1.5) +
  labs(
    y= "% of Nucleosomes", 
    x = "Strands") +
  scale_x_discrete(
    labels = c("Leading L", "Lagging L", "Leading R", "Lagging R")) +
  theme(
    legend.position = "none",
    axis.line = element_line(color = "black", size = 0.5),
    axis.text.x = element_text(color = "black"),
    axis.text.y = element_text(color = "black"),
    plot.title = element_text(hjust = 0.5),
    panel.background = element_blank(),
    plot.background = element_blank()) +
  ggtitle("Total nucleosomes at Right and Left fork")

ggsave("Total_nucleosomes_at_Right_and_left_fork_BC01.svg", plot= p2)

print(p2)
```
F) Calculating nucleosomes in the widom sequences. The analysis will be done first in the total amount of nucleosomes only in the Widom sequences.
```{r}
unique(BC01_Nucleosomes$`Region of Interest`)
head(BC01_Nucleosomes)
```
#Redefined the total of nucleosomes only at the widom sequences
```{r}
T_n_widom <- BC01_Nucleosomes %>% 
  filter(`Region of Interest` != "Others") %>% 
  nrow()
T_n_widom
```

#Filter data, by removing others
```{r}
Widoms <- BC01_Nucleosomes %>% 
  filter(`Region of Interest` !="Others") %>% 
  count(StrandStatus) %>% 
  mutate(percentage = n/T_n_widom *100) %>% 
  mutate(StrandStatus = factor(StrandStatus, levels = c("Lead", "Lag"))) %>% 
  complete(StrandStatus, fill = list(n = 0, percentage = 0))

Widoms

```
#Ploting only the % of nucleosomes at the widom sequences
```{r}
p3 <- ggplot(Widoms, aes(x = StrandStatus, y= percentage, fill = StrandStatus)) +
  geom_bar(
    stat = "identity", 
    color = "black",
    size = 0.25,
    width = 0.8) +
  scale_y_continuous(
    breaks = seq(0, 100, by =10), 
    limits = c(0, 100), 
    expand = c(0,0)) +
  geom_text(
    aes(label = paste0(round(percentage, 1), "%")),
            vjust = -0.5) +
  geom_text(
    aes(label = n),
    vjust = 1.5) +
  labs(
    y= "% of Nucleosomes", 
    x = "Strands") +
  scale_x_discrete(
    labels = c("Leading", "Lagging")) +
  theme(
    legend.position = "none",
    axis.line = element_line(color = "black", size = 0.5),
    axis.text.x = element_text(color = "black"),
    axis.text.y = element_text(color = "black"),
    plot.title = element_text(hjust = 0.5),
    panel.background = element_blank(),
    plot.background = element_blank()) +
  ggtitle("Total nucleosomes at Widom sequences")

ggsave("Total_nucleosomes_at_Widom_sequences_BC01.svg", plot = p3)

print(p3)
```
G) Calculating nucleosomes in the widom sequences.Individual Widom sequences.
```{r}
Widoms_unique <- BC01_Nucleosomes %>% 
  filter(`Region of Interest` !="Others") %>% 
  count(StrandStatus, `Region of Interest`) %>% 
  mutate(percentage = n/T_n_widom *100) %>% 
  mutate(`Region of Interest` = factor(`Region of Interest`, levels = c("Widom626", "Widom612", "Widom605", "Widom601", "Widom603", "Widom607", "Widom613", "Widom618"))) %>% 
  mutate(StrandStatus = factor(StrandStatus, levels = c("Lead", "Lag"))) %>% 
  complete(StrandStatus, `Region of Interest`, fill = list(n = 0, percentage = 0))

Widoms_unique
```
```{r}
p4 <- ggplot(Widoms_unique, aes(x = `Region of Interest`, y = percentage, fill = rev(StrandStatus))) +
  geom_col(
    stat = "identity",
     color = "black",
    size = 0.25,
    position = "dodge",
    width = 0.8) +
  scale_y_continuous(
    breaks = seq(0, 20, by =1), 
    limits = c(0, 20), 
    expand = c(0,0)) +
  geom_text(
    aes(label = paste0(round(percentage, 1), "%")),
            vjust = -0.5,
            position = position_dodge(width = 0.9)) +
   geom_text(
    aes(label = n),
    vjust = 1.5,
    position = position_dodge(width = 0.9)) +
  labs(
    y= "% of Nucleosomes", 
    x = "Strands",
    fill = "Strand") +
  #scale_x_discrete(
   # labels = c("Leading_R", "Lagging_R", "Leading_L", "Lagging_L")) +
  theme(
    #legend.position = "none",
    axis.line = element_line(color = "black", size = 0.5),
    axis.text.x = element_text(color = "black"),
    axis.text.y = element_text(color = "black"),
    plot.title = element_text(hjust = 0.5),
    panel.background = element_blank(),
    plot.background = element_blank()) +
  ggtitle("Nucleosomes at widom sequences")

ggsave("Nucleosomes_at_widom_sequences_BC01.svg", plot = p4)

print(p4)
```



